---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




Create post updated nse with dplyr for version 6.


Intro: Link back to old post. And to the suggestions.
According to [rdocumentation.org](https://www.rdocumentation.org/trends) the dplyr
package is one of the most downloaded packages in R. 

Dplyr is incredibly useful in interactive use, so useful that I want to use 
it in all my packages and programming too. However the programming version  of dplyr
has always been a bit clunky. Not unusable, but not ideal either. 
 HIER UITLEGGEN WAAROM JE DIT WILT. JE WILT FUNCTIE MAKEN DIE DINGEN GAAT DOEN IN 
 DE TOEKOMST, MAAR JE WEET NOG NIET WAT DE NAMEN VAN DE VARIABLEN OF DATAFRAME ZIJN.
 MAAR OMDAT DPLYR DAT WEL WIL WETEN, MOET JE DOOR HOEPELS SPRINGEN

I created a blogpost describing the functions [here]("{{ site.baseurl }}{% link _posts\blog\2016-06-13-NSE_standard_evaluation_dplyr.md %}")

Another option is to use the [wrapr package](https://github.com/WinVector/wrapr). In the
`wrapr` package is a function `let()` that allows you 
to program with standard dplyr functions and 'wrap' some glue code around it to 
make the functions more flexible.

But soon  (May 11th 2017) there will be a dplyr version 0.6.0 a version that eventually will break
your old code where you were using the functions with a underscore: select_(), 
filter_(), etc. There is an excellent description of what to do with your code
and packages in this [compatability guide](http://dplyr.tidyverse.org/articles/compatibility.html#for-users-of-se_-verbs). 

In this update I show what your options were, how to use wrapr,
and how things work in the future. I intent this post to be a reminder that you can
check. Therefore the examples I use are silly and trivial but they are here to explain
the programming part. 


# General idea of programming with dplyr

**interactive use of dplyr**
Dplyr is my main tool to look at data, select sub-parts, check means and all that stuff. 
Whats great is that it requires less typing and faster prototyping. However, the
interactive version of the tools used a special way to work with your input (also called Non Standard Evalation or NSE). In the background the package calls C-code to do
all the manipulations and for normal use it doesn't matter if the data lives on your
drive, or on a database. The verbs all work the same. It's great!

**old style programming of dplyr**
In the previous versions of dplyr (0.5.0) there was an interactive verb, for instance 
`select()` that captured the name of the dataframe and the name of the variables that
you wanted to use and translated that into standard evaluation. That was fed into
the special versions of the verbs `select_()`. So that in essence every call to the
NSE verbs, resulted into a call to the standard verbs. 
To program with dplyr you had to call the standard verbs, to get around the capturing part. 

**glue code and wrapr**
The notation and examples of NSE in dplyr were hard to work with and took quite some effort.
For that reason people have created packages that glue around the edges and do 
substition for you. One of those is the `wrapr` package and I describe the way you can 
use it below. 

**New style programming with dplyr**
It pains me that I have to rewrite some of my code now, but I think Hadley Wickham
and Romain Francois have done an excellent job to make programming with dplyr easier.
The wrote a [vignette](http://dplyr.tidyverse.org/articles/programming.html) that 
describes programming with dplyr very well. So I will only show some examples in the 
verb-section. I borrow heavily from their vignette and refer everyone to that piece of 
knowledge.

The new style programming is based on a new way to capture the environment and the
call, to capture not only the name of the variable but how that variable fits into
the rest of datastructure. "cyl" does not stand alone, but is part of "mtcars".

> We need to quote the input ourselves (not automatically let dplyr it do) and
tell the function not to quote the input. 

So they introduced a set of new verbs, with ... interesting names.

`quo()`, `quos()`, `enquo()`, `quo_name()`, `!!` , `!!!`, `UQ()`, `UQE()`

> quo() is equivalent to quote() and enquo() is equivalent to substitute().)

> quo() works like ": it quotes its input rather than evaluating it.

quo() returns a quosure, which is a special type of formula.

> In dplyr (0.6.0)(and in tidyeval in general) you use !! to say that you want to unquote an input so that it’s evaluated, not quoted. 

```
my_summarise <- function(df, group_by) {
  quo_group_by <- quo(group_by)  # capture the environment
  df %>%
    group_by(!!quo_group_by) %>% # tell dplyr to evaluate here.
    summarise(a = mean(a))
}
```
> By analogy to strings, we don’t want "", instead we want some function that turns an argument into a string. That’s the job of enquo().

quo_name() to convert the input expression to a string

quos(...) converts to a list of formulas


# Specific examples per dplyr verb

## mutate 
> mutate() adds new variables that are functions of existing variables

**The standard interactive way of using mutate**
```{r}
library(dplyr)
mtcars %>% 
        mutate(sum_mpg_wt = mpg + wt) %>% 
        head(3)
```
**The 'old'-version of programming with dplyr**
```
mtcars %>% mutate_(.dots = setNames(list(~mpg+wt), "sum mpg wt"))# give it a name
mtcars %>% mutate_(.dots = list(~mpg+wt)) # if you don't need the name specified
# or
```

the mutate_ function takes the `.dots` argument that describes what you want to do.

```{r}
add_columns_and_name_it <- function(col1, col2, new_col_name) {
    mutate_call <- lazyeval::interp(~ a + b, a = as.name(col1), b = as.name(col2))
    mtcars %>% mutate_(.dots = setNames(list(mutate_call), new_col_name))
}
add_columns_and_name_it("mpg","wt","sum_mpg_wt" ) %>% 
        names()
```

*You first create a call that will be evaluated by mutate_ . the call is first interpreted so that the final and correct names are used by mutate_.*

**wrapr way**
Add the paramters as a list and wrapr's list will substitute the names.

```{r}
library(wrapr)
let(
        alias=list(columnname="sum_mpg_wt",
               col1= "mpg",
               col2="wt"),
        expr={
      mtcars %>% 
        mutate(columnname = col1 + col2)
    }
    ) %>% names()
```

**new style dplyr programming**

```
my_summarise <- function(df, group_by) {
  group_by <- enquo(group_by)
  print(group_by)

  df %>%
    group_by(!!group_by) %>%
    summarise(a = mean(a))
}
```


## select
> select() picks variables based on their names.

Select is used to select columns.

```{r}
mtcars %>% 
        select(mpg,cyl) %>% 
        head(3)
```
**old style dplyr programming**
```{r}
mtcars %>% 
        select_(~mpg, ~cyl) %>% 
        head(3)
```

**wrapr way**
```{r}
let(alias = list(col1 = "mpg", col2 = "cyl"),
    expr = {
            mtcars %>% 
                    select(col1,col2)
    }
)
```
    
    
## filter    
> filter() picks cases based on their values.



## summarise, summarize
> summarise() reduces multiple values down to a single summary.


## arrange
> arrange() changes the ordering of the rows.

## group_by
> apply a grouping with group_by()


Conclusions: new style programming is probably better, still quite difficult.


# Links

- [dplyr 0.6 post - rstudio](https://blog.rstudio.org/2017/04/13/dplyr-0-6-0-coming-soon/)
- [Programming with dplyr - vignette](http://dplyr.tidyverse.org/articles/programming.html)
- [backward compatability help for your old dplyr code](http://dplyr.tidyverse.org/articles/compatibility.html)
