---
title: "plot a day 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
I found this wonderful picture 

INSERT PICTURE HERE

It shows some best practices for displaying information. Now you don't 
have to agree with these practices, but it is rather fun to try to recreate
these plots. I will use the R packages: dplyr, tidyr, ggplot and possibly magrittr. The easiest way to install these packages is with the tidyverse installer:

```
install.packages("tidyverse")
```

http://www.cookbook-r.com/Graphs/ will tell you all about making graphs and changing them

## bar graphs

we will be working with the airquality dataset


```{r}
head(airquality)   # top 6 rows of the dataset 
str(airquality)    # structure
?airquality        # the helppage of this dataset
```

```{r}
library(ggplot2)
ggplot(data = airquality)+ # start a plot layer
        geom_bar(aes(x = Day, y = Ozone),stat = "identity") # bar layer, by day. Y axis Ozone.
```

Normally geom_bar count the number of values in each group. However, I would like to see the Ozone levels on each day. 

Is this really true? more then 300 on day one?

```{r}
airquality %>% 
        filter(Day == 1)
```

Ah there is a measurement every month. So we actually need to plot every month seperate. 
There is a way to create small multiples of a plot. In ggplot this is called faceting. 

```{r}
library(ggplot2)
ggplot(data = airquality)+
        geom_bar(aes(x = Day, y = Ozone),stat = "identity")+
        facet_wrap(facets = ~Month)  # create facets per month. 
```

Now the values are grouped per month. 



* mogelijk per dag van de week, aangeven dat het volgens instructies
van may tot september 1973 is.

* mutate dag en maand plus jaar in nieuwe kolom iso notatie 1973-month-day 
dan weekdays loslaten op datum. of weekend doordeweeks verschillen. 

* opslaan als nieuwe dataframe

* verder werken met nieuwe day of week variabele
* ozone, solar.
