---
title: "Plot a day day 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
I found this wonderful picture 

INSERT PICTURE HERE

It shows some best practices for displaying information. Now you don't 
have to agree with these practices, but it is rather fun to try to recreate
these plots. I will use the R packages: dplyr, tidyr, ggplot and possibly magrittr. The easiest way to install these packages is with the tidyverse installer:

```
install.packages("tidyverse")
```

http://www.cookbook-r.com/Graphs/ will tell you all about making graphs and changing them

## Relations between two variables

We start with a relative simple (in complexity) plot^[I never want to say things are simple, that is extremely frustrating for newcomers, if I do, please correct me!]. Visualizing 
the relation between two variables. 
We start with the build in dataset `airquality`.
A bit of exploring: 
*notice that we don't have to load the dataset.*
*The dataset is actually available and loaded when necessary*

```{r}
head(airquality)   # top 6 rows of the dataset 
str(airquality)    # structure
?airquality        # the helppage of this dataset
```

Then we start with plotting.
I will give the name of the argument en then the variable.
So for instance `ggplot(data = airquality)` note that is not necesarry:
There is a fixed order, see `?ggplot`.
Under **Usage** it says `ggplot(data = NULL, mapping = aes(), ..., environment = parent.frame())`
If you give the arguments in  
the order: data and then mapping, R will assume that you meant it in that order. 
So `ggplot(airquality)` is assumed to mean `ggplot(data = airquality)`. 
Likewise in the `mapping = aes()` part, it says: `aes(x, y, ...)`. 
So I could say aes(OZone, Temp) where Ozone would take the place of x and Temp y. 

```{r plotting two variables}
library(ggplot2) # load ggplot to start plotting
ggplot(data = airquality)+  # first a layer of ggplot
        geom_point(aes(x =Ozone, y= Temp)) # then a layer of points
```

We plot the two variables against each other. 

## Extending the scatterplot to include more then 2 variables.

What if you would like to visualize more variables?

```{r plotting more then 2 variables}
ggplot(data = airquality)+  # first a layer of ggplot
        geom_point(aes(x =Ozone, y= Temp, color =  Solar.R)) # then a layer of points
```

To add more variables we can also change

- transparancy: alpha
- color = colour/color
- size 
- shape

```{r adding size}
ggplot(data = airquality)+  # first a layer of ggplot
        geom_point(aes(x =Ozone, y= Temp, color =  Solar.R, size = Month)) # then a layer 
```

not really helpfull. 

```{r adding size}
ggplot(data = airquality)+  # first a layer of ggplot
        geom_point(aes(x =Ozone, y= Temp, color =  Solar.R, shape = as.factor(Month))) # then a layer 
```

